---
title: "Week 07 Homework"
author: "[Aryaa Rajouria] (mailto:rajouria@uw.edu)"
date: '`r format(Sys.time(), "%Y-%m-%d %H:%M")`'
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
output: 
    html_document:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
urlcolor: blue        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r libs}
# key setups
suppressMessages(library(dplyr))
suppressMessages(library(tidyverse))
suppressMessages(library(knitr))
suppressMessages(library(haven))
suppressMessages(library(magrittr))
#getting rid of popup messages
options(warn=-1)
```

```{r}
#loading data
data <- haven::read_dta("http://staff.washington.edu/phurvitz/csde502_winter_2021/data/AHwave1_v1.dta")
attributes(AHwave1_v1_haven$h1gi1m)$labels
```
```{r}
# use Sys.getenv("TEMP") to get the system temp dir to save the CSV and RDS files
tmpdir <- Sys.getenv("TEMP")
write.csv(x = AHwave1_v1_haven, file = file.path(tmpdir, "AHwave1_v1_haven.csv"), row.names = FALSE)
saveRDS(object = AHwave1_v1_haven, file = file.path(tmpdir, "AHwave1_v1_haven.RDS"))
```

I will first generate a new data frame from the full public Add Health data set. 

```{r}

library(tidyverse)

newdata <- data %>% select(bio_sex, h1gi4, h1fs6, h1fs7) %>% mutate (bio_sex = bio_sex, hisp= h1gi4, dep = h1fs6, tired = h1fs7)

newdata

```

I will then change the labels to be relevant to and informative of my variables of interest.
```{r}
##looking to see if columns have attribute
newdata %>% attributes() %>% map(~head(.))

#converting bio sex to factor
head(newdata$bio_sex)

newdata$bio_sex <- factor(newdata$bio_sex)

bs<- newdata$bio_sex %>% 
    attributes() %>% 
    extract2("labels") %>% 
    names()

newdata$bio_sex <- factor(newdata$bio_sex, 
                                  labels = "biological_sex", 
                                  ordered = TRUE) %>% 
    fct_relevel(rev)
```





```{r}
# extract the labels from the column attribute
#sex <- newdata$bio_sex %>% 
  #  attributes() %>% 
   # extract2("labels") %>% 
    #names()

# create the factor variable
#newdata$bio_sex <- factor(newdata$bio_sex, 
                                 # labels = "bio_sex", 
                                  #ordered = TRUE) %>% 
    #fct_relevel(rev)


# extract the labels from the column attribute
#hispanic <- newdata$h1gi4 %>% 
 #   attributes() %>% 
#    extract2("labels") %>% 
 #   names()

# create the factor variable
#newdata$h1gi4 <- factor(newdata$h1gi4, 
 #                                 labels = #"hispanic", 
                 #                 ordered = TRUE) #%>% 
   # fct_relevel(rev)

# extract the labels from the column attribute
depressed <- newdata$h1fs6 %>% 
    attributes() %>% 
    extract2("labels") %>% 
    names()

# create the factor variable
newdata$h1fs6 <- factor(newdata$h1fs6, 
                                  labels = "depressed", 
                                  ordered = TRUE) %>% 
    fct_relevel(rev)

# extract the labels from the column attribute
#tired <- newdata$h1fs7 %>% 
 #   attributes() %>% 
  #  extract2("labels") %>% 
   # names()

# create the factor variable
#newdata$bio_h1fs7 <- factor(newdata$h1fs7, 
                                 # labels = "tired", 
                              #    ordered = TRUE) #%>% 
#    fct_relevel(rev)
```


```{r}
# factor variable
#tab biosex
(tab_biosex <- newdata %>% 
     group_by(bio_sex) %>% 
     summarise(n = n()))

#factor variable
#tab hispanic or not
(tab_hisp <- newdata %>% 
     group_by(h1gi4) %>% 
     summarise(n = n()))

#tab too depressed
(tab_hisp <- newdata %>% 
     group_by(h1gi6) %>% 
     summarise(n = n()))

#tab too tired to do things 
(tab_hisp <- newdata %>% 
     group_by(h1gi7) %>% 
     summarise(n = n()))
```


```{r}
attributes(newdata)$label <- "National Longitudinal Study of Adolescent to Adult Health (Add Health), 1994-2000 with some variable additions"

attributes(newdata$bio_sex)$label <- "biological sex"
attributes(newdata$bio_sex)$levels <- c("male", "female", "no ans")

attributes(newdata$tired)$label <- "data on feeling tired"
attributes(newdata$tired)$levels <- c("(0) Never/rarely", "(2) A lot of the time", "(3) Most/all the time")

```


```{r}
Sys.getenv("TEMP")
```

Creating tables for the different variables to give us a little more information about the variables of interest.
```{r}
#table w count
round(prop.table(table(newdata$bio_sex)), 2)
#table w percent
round(prop.table(table(newdata$bio_sex)), 2)

#putting it in a nice table
newdata %>% 
    group_by(bio_sex) %>% 
    summarise(n = n()) %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    knitr::kable() 
```
```{r}
#table w count
round(prop.table(table(newdata$h1gi4)), 2)
#table w percent
round(prop.table(table(newdata$h1gi4)), 2)

#putting it in a nice table
newdata %>% 
    group_by(h1gi4) %>% 
    summarise(n = n()) %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    knitr::kable() 

```

```{r}
#table w count
round(prop.table(table(newdata$h1fs6)), 2)
#table w percent
round(prop.table(table(newdata$h1fs6)), 2)

#putting it in a nice table
newdata %>% 
    group_by(h1fs6) %>% 
    summarise(n = n()) %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    knitr::kable() 

```

```{r}
#table w count
round(prop.table(table(newdata$h1fs7)), 2)
#table w percent
round(prop.table(table(newdata$h1fs7)), 2)

#putting it in a nice table
newdata %>% 
    group_by(h1fs7) %>% 
    summarise(n = n()) %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    knitr::kable() 
```

Now I tabulate biological sex with the variable depression. 
```{r}
newdata %>% 
    group_by(bio_sex,
             h1fs7) %>% 
    summarise(n = n(), .groups = "drop_last") %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    kable() 
```

Finally, we graph biological sex with depression. 

```{r}
#using ggplot to graph depression and biological sex
boxplot <- ggplot(data=newdata, aes(x=bio_sex, y=h1fs7)) +
              geom_boxplot()

boxplot #called to view boxplot 
```
