---
title: "Week 05 Homework"
author: "[Aryaa Rajouria] (mailto:rajouria@uw.edu)"
date: '`r format(Sys.time(), "%Y-%m-%d %H:%M")`'
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
output: 
    html_document:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
urlcolor: blue        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r libs}
# key setups
suppressMessages(library(dplyr))
suppressMessages(library(tidyverse))
suppressMessages(library(knitr))
suppressMessages(library(haven))

options(warn=-1)
```

## Setting up the dataset 

Below, I generate the dataset for the pain of graduate school experienced by graduate students (gradpain). 
```{r data}
gradpain <- c(rep(0,11), rep(1,1), rep(2,2), rep(3,6), rep(4,8), rep(5,10),
 rep(6,8), rep(7,30), rep(8,10), rep(9,6), rep(10,2))

barplot <-barplot(table(gradpain), las=1, ylab="Frequency", xlab="Pain Scale",
 main="How painful is graduate school?")

table(gradpain)
summary(gradpain)
```

```{r}
#pain <- table (gradpain)
#pain

#as.data.frame(table(pain))
#table(gradpain)

#count(gradpain[1:94])

#count_pain <- function(gradpain){
  #df <- data.frame(gradpain = gradpain[1:length(gradpain)])
  #count(df[1:94])
#}

#count_pain
```

## Box 1


```{r boxone}
fun_sam <- function(x, d.vec= gradpain, n= length(gradpain)){
    s <- sample(x= gradpain, size = 1000, replace = TRUE)
  return(list(mean(d.vec), n))
}

boxone <- fun_sam(gradpain)
```

As per the function fun_sam, the mean and sample size of the function is `r boxone`.

## Box 2

```{r}
fun_samp <- function (d.vec = gradpain, n = 100){
  new_s <- sample (x = d.vec, size=n, replace= TRUE)
  list(len_ = length(samp), mean_= (samp))
}
```

```{r}

#taking sample from d.vec, of 100 n w/ replacement 
fun_samp <- function(d.vec = gradpain, n = 100){
    samp <- sample(x = d.vec, size = n, replace = TRUE)
    list(vec.len = length(samp),
         vec.mean = mean(samp))
}

# create an empty vector to store results
x <- NULL

# setting seed for replicabiity 
set.seed(7)

# run 100x loop: esp the fun_sample, in a list form 
for(i in 1:100){
    m <- fun_samp(d.vec = gradpain, n = length(gradpain))$vec.mean
    # build up the vector
    x <- c(x, m)
}
```

As per the function, the mean and standard deviation for some of the samples are `r head(x)` of the randomly sample means.

## Box 3

```{r boxthree}
set.seed(7)
function6 <- function(d.vec= gradpain, n= length(gradpain), num.samples=100){
  L <- NULL 
for (i in 1:num.samples){
  s <- sample(x= d.vec, size = n, replace = TRUE)
  L <- c(x, mean(s))
  meanmeans <- mean(L)
  sdf <- sd(L)
  ci_95 <- Rmisc::CI(x = L, ci =0.95)
  return(list(n, num.samples, L,  meanmeans, sdf, ci_95))
}}

numbers <- function6(as.numeric(gradpain))
numbers
```
5. What does your function return for the mean of means?

The function returns `r numbers$meanmeans` for the mean of means. 

6. What does your function return for the standard deviation of means?

The function returns `r numbers$sdf` for the standard deviation of means. 

7. What does your function return for the 95% confidence interval around the mean?

The function returns `r numbers$ci_95` for the 95% confidence interval around the mean. 