---
title: "Week 08 Homework"
author: "[Aryaa Rajouria] (mailto:rajouria@uw.edu)"
date: '`r format(Sys.time(), "%Y-%m-%d %H:%M")`'
header-includes: #allows you to add in your own Latex packages
- \usepackage{float} #use the 'float' package
- \floatplacement{figure}{H} #make every figure with caption = h
output: 
    html_document:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
urlcolor: blue        
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```


```{r libs}
# key setups
suppressMessages(library(dplyr))
suppressMessages(library(tidyverse))
suppressMessages(library(knitr))
suppressMessages(library(haven))
suppressMessages(library(magrittr))
suppressMessages(library(kableExtra))
library(pdftools)
library(curl)
library(ggplot2)
library(naniar)
#getting rid of popup messages
options(warn=-1)
```

```{r}
library(captioner)

figure_nums <- captioner(prefix = "Figure")
table_nums <- captioner(prefix = "Table")
```

#
Imagine a new variable: multirace, using the following value definitions:

	1 = one race, White
	2 = one race, not White
	3 = two races, includes White
	4 = two races, both non-White
	5 = three or more races, includes White
	6 = three or more races, all non-White
	9 = any race missing (White, Black/African American, American Indian, Asian, other)


## 
__Fill in the codes for the hypothetical cases below (`r table_nums(name = "t01", display = "cite")`).__

`r table_nums(name = "t01", caption = "A hypothetical data set")`


| white| black| AI| asian| raceother|multirace |
|-----:|-----:|--:|-----:|---------:|---------:|
|     1|     0|  0|     0|         0|        0 |
|     0|     1|  0|     0|         0|        0 |
|     1|     0|  0|     1|         0|        1 |
|     0|     1|  1|     0|         0|        1 |
|     1|     1|  0|     1|         0|        1 |
|     0|     1|  0|     0|         1|        1 |
|     0|     1|  1|     0|         1|        1 |
|     1|     0|  1|     0|         0|        1 |
|     1|     1|  1|     0|         1|        1 |
|     6|     1|  8|     1|         6|        7 |	


## 
__Using this data frame (code below), report how many cases checked more than one race.__ Use R code to make this calculation and use inline expressions.

```
dat <- 
structure(
    list(
        white = c(1L, 0L, 1L, 0L, 1L, 0L, 0L, 1L, 1L, 6L),
        black = c(0L, 1L, 0L, 1L, 1L, 1L, 1L, 0L, 1L, 1L), 
        AI = c(0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 1L, 8L), 
        asian = c(0L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 1L), 
        raceother = c(0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 6L), 
        multirace = c(0L, 0L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 7L),
    ), 
    class = "data.frame", 
    row.names = c(NA, -10L)
)
```
```{r}
dat <- 
structure(
    list(
        white = c(1L, 0L, 1L, 0L, 1L, 0L, 0L, 1L, 1L, 6L),
        black = c(0L, 1L, 0L, 1L, 1L, 1L, 1L, 0L, 1L, 1L), 
        AI = c(0L, 0L, 0L, 1L, 0L, 0L, 1L, 1L, 1L, 8L), 
        asian = c(0L, 0L, 1L, 0L, 1L, 0L, 0L, 0L, 0L, 1L), 
        raceother = c(0L, 0L, 0L, 0L, 0L, 1L, 1L, 0L, 1L, 6L), 
        multirace = c(0L, 0L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 7L)
    ), 
    class = "data.frame", 
    row.names = c(NA, -10L)
)
```


##
__Write R code to create the multirace variable, using the data set [AHwave1_v3.rds](http://staff.washington.edu/phurvitz/csde502_winter_2021/data/AHwave1_v3.rds).__ _Hint: You may want to create another variable, numrace, that counts the number of races._ Use [`download_file()`](https://stat.ethz.ch/R-manual/R-devel/library/utils/html/download.file.html) and `Sys.getenv("TEMP")` to download the file to your system's `TEMP` directory to avoid local file system dependencies.


```{r}
#loading data
v3 <- "http://staff.washington.edu/phurvitz/csde502_winter_2021/data/AHwave1_v3.rds"
  
download.file(v3, "file.RDS")
# read into R
readRDS("file.RDS")

# Or similarly
curl::curl_download(v3, "file.RDS")
# read into R
readRDS("file.RDS")

#saving to temp 
getting_data <- function(v3) {
  
  tempFile_location<- tempfile()
  download.file(v3, tempFile_location)
  b <- readRDS(tempFile_location)
  file.remove(tempFile_location)
  b
}

b <- getting_data(v3)
```

```{r}
#creating a temp directory so that you don't have to depend on local file system dependencies
Sys.getenv("TEMP")
Sys.getenv("TMPDIR")

```


## 
__Label the multirace variable as well as its values using attribute labels.__.

```{r}
#labeling variable using attribute labels
attributes(dat$multirace)$label <- "multirace"
attributes(dat$multirace)

# label for health
attributes(dat$multirace)$label <- "Multirace data"
# values for health
attributes(dat$multirace)$levels <- c("(1) individual is multiracial", "(0) individual is not multiracial", "(7) total number of individuals that are multiracial")
```


## 
__Include below a contingency table of the multirace variable. Make sure that the values are labelled so the table is readable, and also include any missing values.__

```{r}
#table w count
round(prop.table(table(dat$multirace)), 2)
#table w percent
round(prop.table(table(dat$multirace)), 2)

#putting it in a nice table
dat %>% 
    group_by(multirace) %>% 
    summarise(n = n()) %>% 
    mutate(`%` = n / sum(n) * 100) %>% 
    mutate(`%` = `%` %>% round(1)) %>% 
    knitr::kable(caption = "Multirace Variable") %>% 
    kable_styling(full_width = FALSE, position = "left")
```